---
title: "Supermarket Sales"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(ggplot2)
library(dplyr)
getwd()
setwd("/Users/robertvargas/Documents/Projects/R/Supermarket sales")
data<- read.csv("supermarket_sales - Sheet1.csv")

##context - create quarterly report
## give context of sales
print("total sales for Q1 were...")
total_sales = format(round(sum(data$Total),0), big.mark = ",", scientific = FALSE)
#pie chart - sales efrom each branch
branch_sales<- aggregate(data$Total, by = list(data$Branch), FUN = sum)
  colnames(branch_sales)<- c("Branch", "Q1_Sales")
  branch_sales$Q1_Sales<- round(branch_sales$Q1_Sales,0)
  branch_sales$prop<- round(branch_sales$Q1_Sales/ sum(branch_sales$Q1_Sales)*100,0)
branch_sales<- branch_sales %>% arrange(desc(Branch)) %>% mutate(ypos = cumsum(prop) - 0.5*prop)
 
sales_pie_chart <- ggplot(branch_sales, aes( x = "", y = prop, fill = Branch)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar("y", start = 0 ) +
  geom_text(aes(y = ypos, label = paste(prop,"%")), color = "white") +
  scale_fill_manual(values = c("#0073C2FF", "#EFC000FF", "#CD534CFF")) +
  ggtitle("Branch Sales as % of Total") +
  theme_void()

#bar plot, sales by for each product line
sales_by_product_line = aggregate(data$Total, by = list(data$Product.line), FUN = sum)
  colnames(sales_by_product_line)<- c("Product", "Q1_Sales")
  sales_by_product_line$Q1_Sales = round(sales_by_product_line$Q1_Sales,0)

sales_barplot<- ggplot(data = sales_by_product_line, aes( x = Product, y = Q1_Sales)) +
  geom_bar(stat = "identity", fill = "#0073C2FF", color = "black" ) +
  labs(y = "Sales ($)" ) +
  ggtitle("Product Line Sales") +
  coord_flip( )
  
## double bar graph, member/gender
customer_freq_table<- table(data$Gender, data$Customer.type)
customer_freq_table<- as.data.frame(customer_freq_table)
colnames(customer_freq_table) <- c("Gender", "Membership", "Freq")


customer_barplot <- ggplot(data = customer_freq_table, aes(x = Gender, y = Freq, condition = Gender, fill = Gender)) +
  facet_wrap(~ Membership) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_manual(values = c("#0073C2FF", "#EFC000FF")) +
  ggtitle("Customer Analytics") +
  labs(x = "", y = "") +
  theme_minimal()

#overall shopping experience
mean_rating = mean(data$Rating)
median_rating = median(data$Rating)
customer_hist <- ggplot(data = data, aes(x = data$Rating)) +
  geom_histogram(binwidth = .5, color = "black", fill = "#0073C2FF") +
  ggtitle("Customer Ratings Frequency") +
  labs( x = "Rating Scale from 1 - 10", y = "")
  theme_minimal()

#exploratory data analysis
#sales forecast per week, day
data$Date<- as.Date(data$Date, format = "%m/%d/%Y")
sales_forecast<- aggregate(data$Total, by = list(data$Date), FUN = sum)
  sales_forecast<- as.data.frame(sales_forecast)
  colnames(sales_forecast) <- c("Date","Total_Sales")
  sales_forecast$Total_Sales<- round(sales_forecast$Total_Sales, 0)
  
sales_forecast_graph <- ggplot(sales_forecast, aes( x = Date, y = Total_Sales)) +
  geom_line( color = "#CD534CFF") +
  ggtitle("Sales Forecast for Quarter 1") +
  labs( x = "Month")
  ##break into weeks
sales_forecast$Week<- as.Date(cut(sales_forecast$Date, breaks = "week", start.on.monday = FALSE))
weekly_sales_graph <- ggplot(data = sales_forecast, aes(x = Week, y = Total_Sales)) +
  stat_summary(fun.y = sum, geom = "bar", fill = "#CD534CFF", color = "black") +
  ggtitle("Sales Forecast by Week")

# timeline of no of purchases
data$Day<- weekdays(data$Date)
data$Day<- as.factor(data$Day)
total_order_by_days<- aggregate(data$Invoice.ID, by = list(data$Day), FUN = length)
  total_order_by_days<- as.data.frame(total_order_by_days)
  colnames(total_order_by_days)<- c("Day", "Total_Orders")

total_orders_per_day <- ggplot(total_order_by_days, aes(x = Day, y = Total_Orders)) +
  geom_bar(stat = "identity", color = "black", fill = "#EFC000FF") +
  ggtitle("Total Orders per Day") +
  labs (y = "No. of Orders")

avg_sale_pmttype<- aggregate(data$Total,by = list(data$Payment), FUN = mean)
  avg_sale_pmttype<- as.data.frame(avg_sale_pmttype)
  colnames(avg_sale_pmttype)<- c("Payment_type", "Avg_Trans_Amount")
  avg_sale_pmttype$Avg_Trans_Amount<- paste("$",round(avg_sale_pmttype$Avg_Trans_Amount,0))

##total sales per hour
data$Hour<- substr(data$Time,1,2)
total_orders_hour<- aggregate(data$Invoice.ID, by = list(data$Hour), FUN = length)
total_orders_hour<- as.data.frame(total_orders_hour)
colnames(total_orders_hour)<- c("Hour", "Total_Orders")

hour_barplot<- ggplot(total_orders_hour, aes ( x = Hour, y = Total_Orders )) +
  geom_bar(stat = "identity", color = "black", fill = "#EFC000FF") +
  ggtitle("Total Orders per Hour") +
  labs( y = "No. of Orders")

## sales per month 
data$Month<- format(data$Date, "%m")
sales_per_month<- aggregate(data$Total, by = list(data$Month), FUN = sum)
  sales_per_month<- as.data.frame(sales_per_month)
  colnames(sales_per_month) <- c("Month", "Total_Sales")
  sales_per_month$Total_Sales<- format(round(sales_per_month$Total_Sales,0), big.mark = ",", scientific = FALSE)

## customer aggregate
  customer_info_sumsales<- aggregate(data$Total, by = list( data$Customer.type, data$Gender, data$Product.line, data$Day), FUN = sum)
  customer_info_sumsales<- as.data.frame(customer_info_sumsales)
  colnames(customer_info_sumsales)<- c("Customer","Gender", "Product", "Day", "Sales")
  customer_info_sumsales$Sales<- round(customer_info_sumsales$Sales,0)
  
  
  customer_info_nosales<- aggregate(data$Total, by = list( data$Customer.type, data$Gender, data$Product.line, data$Day), FUN = length)
  customer_info_nosales<- as.data.frame(customer_info_nosales)
  colnames(customer_info_nosales)<- c("Customer","Gender", "Product", "Day", "Orders")

  
```

Sales Summary
===
Row 
-----------------------------------------------------------------------

### Chart A
```{r}
valueBox(value = total_sales , caption = "Total Sales for Q1 2019", color = "green")
```

Row
-----------------------------------------------------------------------

### Branch Performance
```{r}
sales_pie_chart
```

### Product Performance
```{r}
sales_barplot
```


Forecasts
===

Row 
-----------------------------------------------------------------------
### Chart
```{r}
valueBox(sales_per_month[1,2], caption = "January", color = "green")
```

### Chart
```{r}
valueBox(sales_per_month[2,2], caption = "February", color = "green")
```

### Chart
```{r}
valueBox(sales_per_month[3,2], caption = "March", color = "green")

```




Row 
-----------------------------------------------------------------------

### Monthly Performance
```{r}
sales_forecast_graph 
```

### Weekly Performance
```{r}
weekly_sales_graph
```


Customer Analytics
===
Row 
-----------------------------------------------------------------------

### Average Customer Rating 
```{r}
gauge(round(mean_rating,2), min = 0 , max = 10)

```

Row
-----------------------------------------------------------------------

### Demographics
```{r}
customer_barplot
```

### Quality of Shopping Experience
```{r}
customer_hist
```

Shopping Behavior
===

Row 
-----------------------------------------------------------------------
### Daily Performance
```{r}
total_orders_per_day
```

### Hourly Performance
```{r}
hour_barplot
```

Row 
-----------------------------------------------------------------------
### Chart C
```{r}
valueBox(value = avg_sale_pmttype[1,2] , caption = "Average Cash Transaction", color = "blue")

```

### Chart D
```{r}
valueBox(value = avg_sale_pmttype[2,2] , caption = "Average Credit Transaction", color = "red")
```

### Chart E
```{r}
valueBox(value = avg_sale_pmttype[3,2] , caption = "Average Ewallet Transaction", color = "green")
```

Final Results
===

Row 
-----------------------------------------------------------------------
### Sales Activity
- Branch B earned the most revenue but each branch generated approximately 33-34% of total sales.
- The best selling products were "Food and beverages", "Sports and travel" was a close second.
- Sales were highest in January, despite sales being the lowest during the first week of January.
- Sales momentum dipped slightly during mid-February. 

Row 
-----------------------------------------------------------------------
### Demographics
- The average customer experience rating was below 7, which might be concerning. The company recieved no ratings below a 4. 
- Most of the customers in quarter 3 were either female members or non-member males.
- Most orders were made on a Saturday or a Tuesday. Monday generally has the lowest frequency of orders. 


Row 
-----------------------------------------------------------------------
### Other
- Most sales were made during 7pm (19) and the least amount of sales occurred during the hours between 4-5pm (16-17) and an hour from closing (20).
- Method of payment has very little influence on the customers transaction amount, as the average transaction amount for each payment type are consistent within $7.

Column 
-----------------------------------------------------------------------

### The customer who made the most orders can be categorized by the following demographics:
```{r}
head(customer_info_nosales[order(-customer_info_nosales$Orders),],1)
```

### The customer who generated the most sales can by categorized by the following demographics:
```{r}
head(customer_info_sumsales[order(-customer_info_sumsales$Sales),],1)
```

